Hartree-Fock Stability

### 2016-11-08 presentations/GroupMeeting_Fall2016.tex
Presentation updates
Added more figures and finishe groupmeeting

\usetheme{metropolis}
\useoutertheme{metropolis}
\useinnertheme{metropolis}
\usefonttheme{metropolis}
%\usecolortheme{seahorse}   % or your preferred color theme
\usepackage{appendixnumberbeamer}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{xspace}
\usepackage[justification=centering]{caption}
\usepackage{braket}
\usepackage{amsmath}

@@

{%
	\setbeamertemplate{frame footer}{Thouless, D. J. The Quantum Mechanics of Many-Body Systems, 2nd ed.; Dover: 1972. (p. 114)}
\begin{frame}{Hartree-Fock Stability}
	\begin{itemize}[<+->]
		\item Solving the HF equations guarantees only that the energy is \alert{stationary}. The first order variation is therefore 0. 
		\item We need to know if this is indeed a minimum.
		\item We can determine this if we inspect the second order variation in the energy.
		\item Thouless$^1$ showed a physically motivated derivation using Time-Dependent Hartree-Fock theory (TDHF). 
	\end{itemize}
\end{frame}

@@

\begin{frame}{Hartree-Fock Stability Conditions}
	\begin{itemize}[<+->]
		\item{The RPA frequencies represent electronic oscillations about the ground state}
		\item{Real frequencies are stable, imaginary frequencies are unstable}
		\item{The RPA frequencies are imaginary when the \alert{Orbital Hessian} (aka stability matrix, electronic Hessian) eigenvalues are negative
			\begin{eqnarray}
				\begin{bmatrix}
					\bf{A}   & \bf{B}   \\
					\bf{B}^* & \bf{A}^* \\
				\end{bmatrix}
				\begin{bmatrix}  \bf{X} \\ \bf{Y}  \end{bmatrix}
				= \omega \begin{bmatrix}  \bf{X} \\ \bf{Y}  \end{bmatrix}
			\end{eqnarray}
		}
		\item{Where}
		\begin{eqnarray}
			A_{ia,jb} &=& \braket{{}_i^a|H-\braket{0|H|0}|{}_j^b} = \left(\epsilon_a - \epsilon_i \right) \delta_{ij}\delta_{ab} + \braket{aj||ib}
			\nonumber \\
			B_{ia,jb} &=& \braket{{}_{ij}^{ab}|H-\braket{0|H|0}|0} = \braket{ab||ij}.
		\end{eqnarray}
	\end{itemize}
\end{frame}

\begin{frame}{Hartree-Fock Stability Conditions}
	\begin{itemize}
		\item[]{The eigenvalue equation can be factorized depending on symmetry.}
		\item[]{\includegraphics[width=\linewidth]{../images/seeger_stability.pdf} }
	\end{itemize}
\end{frame}

@@

\begin{frame}{Brief Overview}
	\begin{itemize}
		\item{The discretized solutions are given by,
		\begin{eqnarray}\label{eq:hf_orb_energy}
			\epsilon_{\vec{k}}=
				\frac{\hbar^2 k^2}{2m} - \sum\limits_{\vec{k}}^{|\vec{k}|< k_f}
				\braket{\vec{k}, \vec{k}' |\vec{k}', \vec{k}}
		\end{eqnarray}
		}
		\item{Where the two electron integral is given by
		\begin{align}
			\braket{\vec{k}, \vec{k}'|\vec{k}'',\vec{k}'''}\stackrel{\text{2D, 3D}}{=}&
				\begin{cases} 
				\frac{\pi}{V}\frac{2^{D-1}}{|\vec{k}-\vec{k}''|^{D-1}} 
				& \vec{k}''' = \vec{k}+\vec{k}'-\vec{k}'' \nonumber\\
				0 
				& \text{else}
				\end{cases}
			\nonumber \\
			\braket{k, k'|k'',k'''}\stackrel{\text{1D}}{=}&
				\begin{cases} 
				e^{|k-k''|^2a^2}\text{Ei}(-|k-k''|^2a^2)\text{ ;}
				& k''' = k+k'-k'' \\
				0\text{ ;} 
				& \text{else}
				\end{cases}
		\end{align}
		}
	\end{itemize}
\end{frame}

@@

 2. Li, R.-C.; Zhang, L.-H. Convergence of Block Lanczos Method for Eigenvalue Clusters; 2013.}
\begin{frame}{Convergence Properties}
	\begin{itemize}[<+->]
		\item{The convergence of these subspace algorithms depends on:}
		\begin{itemize}
			\item{\alert{Number of eigenvalues} requested}
			\item{The \alert{block size} of the algorithm}
			\item{The \alert{density} of the eigenvalue spectrum}
			\item{The \alert{initial guess} eigenvectors}
			\item{The \alert{condition number} of the matrix}
			\item{The \alert{preconditioner} (Davidson only)}
		\end{itemize}
		\item{They can \alert{converge to an incorrect answer} if:}
		\begin{itemize}
			\item{The block size is too small compared to degeneracy}
			\item{The approximate eigenvectors become non-orthogonal}
		\end{itemize}
		\item{My recommendation for guess eigenvectors is
		\begin{eqnarray}
			v_j^{(i)} = normalize\left(\frac{1}{|A_{ii} - A_{jj}| + 1} \right).
		\end{eqnarray}
		}
	\end{itemize}
\end{frame}

{%
\setbeamertemplate{frame footer}{Hern, V.; Tom, A.; Vidal, V. SLEPc Technical Reports, 2007.}
\begin{frame}{Orthogonalization}
	\begin{itemize}
		\item{The condition number, $\kappa$, is bound from below by
			\begin{eqnarray}
				\kappa \geq \frac{Max(\lambda_i)}{Min(\lambda_j)}
			\end{eqnarray}
		}
		\item{The Gram-Schmidt procedure has numerical issues, 
			\begin{eqnarray}
				||\mathbf{I} - \mathbf{Q^TQ}|| \leq \frac{\alpha\kappa^2}{1-\beta\kappa^2}.	
			\end{eqnarray}
		}
		\item{Modified Gram-Schmidt is better, but not perfect,
			\begin{eqnarray}
				||\mathbf{I} - \mathbf{Q^TQ}|| \leq \frac{\gamma\kappa}{1-\eta\kappa}	
			\end{eqnarray}
		}
		\item{May need multiple orthogonalization steps}
	\end{itemize}

@@

@@

\begin{frame}{Davidson Scaling}
	\begin{columns}[c] % align columns
		\begin{column}{.4\textwidth}
			\begin{itemize}
				\item {Davidson is Asymptotically quadratic.}
				\item {Full diagonalization is almost cubic.}
			\end{itemize}				
		\end{column}
		\hfill
		\begin{column}{0.6\textwidth}
		    \begin{overprint}
			    \onslide<1>\includegraphics[width=\linewidth]{../images/dav_vs_exact_scaling.eps}

@@

\begin{frame}{Efficacy of Davidson's Algorithm}
	\begin{columns}[c] % align columns
		\begin{column}{.4\textwidth}
			\begin{itemize}
				\item {Scales as $N_{exc} \times  N_{occ}$ due to momentum conservation}
			\end{itemize}				
		\end{column}
		\hfill
		\begin{column}{0.6\textwidth}
		    \begin{overprint}
			    \onslide<1>\includegraphics[width=\linewidth]{../images/dav_vs_exact.eps}

@@

{%
\setbeamertemplate{frame footer}{Bernu, B.; Delyon, F.; Holzmann, M.; Baguet, L. Phys. Rev. B 2011, 84 (11), 115115.}
\begin{frame}{Dependence on r${}_s$}
	\begin{columns}[c] % align columns
		\begin{column}{.5\textwidth}
			\includegraphics[width=0.95\linewidth]{../images/stab_rs.eps}
			\\
			My Result
		\end{column}
		\hfill
		\begin{column}{0.5\textwidth}
			\includegraphics[width=\linewidth]{../images/Bernu_2dEnergyGain.pdf}
			\\
			Published Result. Red = Square Lattice, Blue = Triangular Lattice
		\end{column}	
	\end{columns}
\end{frame}

@@

{%
\setbeamertemplate{frame footer}{Supady, A.; Blum, V.; Baldauf, C. J. Chem. Inf. Model. 2015, 55 (11), 2338–2348.}
\begin{frame}{Concluding Remarks}
	\begin{itemize}[<+->]
		\item {Finding all the low energy conformers for a molecule is hard}
		\item {In order to get accurate energies/structures, computationally expensive methods should be employed}
		\item {These take time, so we want to minimize how many of these we do}
		\item {The Genetic Algorithm provides a framework for a refined global search}
		\item {It shines when asked to find a host of low energy solutions}
		\item {GA wrapper can be interfaced with a variety of electronic structure packages(NWChem, ORCA) and is available under the GNU Lesser General Public License at \url{https://github.com/adrianasupady/fafoom}}

	\end{itemize}
\end{frame}
}

\begin{frame}[standout]
  Questions?
\end{frame}

@@

\begin{frame}[fragile]{Backup slide}
	Empty Backup
\end{frame}

\end{document}

>>>>

\usetheme{metropolis}
\useoutertheme{metropolis}
\useinnertheme{metropolis}
\usefonttheme{metropolis}
%\usecolortheme{seahorse}   % or your preferred color theme
\usepackage{appendixnumberbeamer}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{xspace}
\usepackage[justification=centering]{caption}
\usepackage{braket}
\usepackage{amsmath}

@@

{%
	\setbeamertemplate{frame footer}{Thouless, D. J. The Quantum Mechanics of Many-Body Systems, 2nd ed.; Dover: 1972. (p. 114)}
\begin{frame}{Hartree-Fock Stability}
	\begin{itemize}[<+->]
		\item Solving the HF equations guarantees only that the energy is \alert{stationary}. The first order variation is therefore 0. 
		\item We need to know if this is indeed a minimum.
		\item We can determine this if we inspect the \alert{second order variation} in the energy.
		\item Thouless$^1$ showed a physically motivated derivation using Time-Dependent Hartree-Fock theory (TDHF). 
	\end{itemize}
\end{frame}

@@

\begin{frame}{Hartree-Fock Stability Conditions}
	\begin{itemize}[<+->]
		\item{The RPA frequencies represent electronic oscillations about the ground state}
		\item{Real frequencies are stable, imaginary frequencies are unstable}
		\item{The RPA frequencies are imaginary when the \alert{Orbital Hessian} (aka stability matrix, electronic Hessian) eigenvalues are \alert{negative}
			\begin{eqnarray}
				\begin{bmatrix}
					\bf{A}   & \bf{B}   \\
					\bf{B}^* & \bf{A}^* \\
				\end{bmatrix}
				\begin{bmatrix}  \bf{X} \\ \bf{Y}  \end{bmatrix}
				= \omega \begin{bmatrix}  \bf{X} \\ \bf{Y}  \end{bmatrix}
			\end{eqnarray}
		}
		\item{Where}
		\begin{eqnarray}
			A_{ia,jb} &=& \braket{{}_i^a|H-E_0|{}_j^b} = \left(\epsilon_a - \epsilon_i \right) \delta_{ij}\delta_{ab} + \braket{aj||ib}
			\nonumber \\
			B_{ia,jb} &=& \braket{{}_{ij}^{ab}|H-E_0|0} = \braket{ab||ij}.
		\end{eqnarray}
	\end{itemize}
\end{frame}

{%
	\setbeamertemplate{frame footer}{1) Seeger, R.; Pople, J. A. J. Chem. Phys. 1977, 66 (7), 3045.}
\begin{frame}{Hartree-Fock Stability Conditions}
	\begin{itemize}
		\item[]{The eigenvalue equation can be factorized depending on symmetry.} \\~\\
		\item[]{
			\resizebox{0.9\textwidth}{!}{%
			\begin{tabular}{r@{\qquad}cccccc}
				  \multirow{1}{*}{Solution Type} & \multicolumn{6}{c}{Space Type}  \\
				  \cmidrule{2-7}
				  & Real RHF & Complex RHF & Real UHF & Complex UHF & Real GHF & Complex GHF \\
				  \midrule
				  Real RHF    & ${}^1\mathbf{A}' + {}^1\mathbf{B}'$ & ${}^1\mathbf{A}' - {}^1\mathbf{B}'$ & ${}^3\mathbf{A}' + {}^3\mathbf{B}'$ & & & \\
				  Complex RHF & - & ${}^1\mathbf{H}'$ & - & \alert{${}^3\mathbf{H}'$} & - & \\
				  Real UHF    & - & - & $\mathbf{A}' + \mathbf{B}'$ & $\mathbf{A}' - \mathbf{B}'$ & $\mathbf{A}'' + \mathbf{B}''$ &  \\
				  Complex UHF & - & - & - &  $\mathbf{H}'$ & - & $\mathbf{H}'$ \\
				  Real GHF    & - & - & - & - & $\mathbf{A} - \mathbf{B}$ &  $\mathbf{A} - \mathbf{B}$  \\
				  Complex GHF & - & - & - & - & - & $\mathbf{H}$ \\
				  \bottomrule
			\end{tabular}}
		\item[] \tiny Table reproduced from Seeger \& Pople$^1$
		}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Brief Overview}
	\begin{itemize}
		\item{The discretized solutions are given by,
		\begin{eqnarray}\label{eq:hf_orb_energy}
			\epsilon_{\vec{k}}=
				\frac{\hbar^2 k^2}{2m} - \sum\limits_{\vec{k}'}^{|\vec{k}'|< k_f}
				\braket{\vec{k}, \vec{k}' |\vec{k}', \vec{k}}
		\end{eqnarray}
		}
		\item{Where the two electron integral is given by
		\begin{align}
			\braket{\vec{k}, \vec{k}'|\vec{k}'',\vec{k}'''}\stackrel{\text{2D, 3D}}{=}&
				\begin{cases} 
				\frac{\pi}{V}\frac{2^{D-1}}{|\vec{k}-\vec{k}''|^{D-1}} 
				& \vec{k}''' = \vec{k}+\vec{k}'-\vec{k}'' \nonumber\\
				0 
				& \text{else}
				\end{cases}
			\nonumber \\
			\braket{k, k'|k'',k'''}\stackrel{\text{1D}}{=}&
				\begin{cases} 
				e^{|k-k''|^2a^2}\text{Ei}(-|k-k''|^2a^2)\text{ ;}
				& k''' = k+k'-k'' \\
				0\text{ ;} 
				& \text{else}
				\end{cases}
		\end{align}
		}
	\end{itemize}
\end{frame}

@@

 2. Li, R.-C.; Zhang, L.-H. Convergence of Block Lanczos Method for Eigenvalue Clusters; 2013.}
\begin{frame}{Convergence Properties}
	\begin{itemize}[<+->]
		\item{The convergence of these subspace algorithms depends on:}
		\begin{itemize}
			\item{\alert{Number of eigenvalues} requested}
			\item{The \alert{block size} of the algorithm}
			\item{The \alert{density} of the eigenvalue spectrum}
			\item{The \alert{initial guess} eigenvectors}
			\item{The \alert{condition number} of the eigenvector matrix}
			\item{The \alert{preconditioner} (Davidson only)}
		\end{itemize}
		\item{They can \alert{converge to an incorrect answer} if:}
		\begin{itemize}
			\item{The block size is too small compared to degeneracy}
			\item{The approximate eigenvectors become non-orthogonal}
		\end{itemize}
		\item{My recommendation for guess eigenvectors is
		\begin{eqnarray}
			v_j^{(i)} = normalize\left(\frac{1}{|A_{ii} - A_{jj}| + 1} \right).
		\end{eqnarray}
		}
	\end{itemize}
\end{frame}

{%
\setbeamertemplate{frame footer}{Hern, V.; Tom, A.; Vidal, V. SLEPc Technical Reports, 2007.}
\begin{frame}{Orthogonalization}
	\begin{itemize}
		\item{The condition number, $\kappa$, is bound from below by
			\begin{eqnarray}
				\kappa \geq \frac{Max(A_{ii})}{Min(A_{jj})}
			\end{eqnarray}
		}
		\item{The Gram-Schmidt procedure has numerical issues, 
			\begin{eqnarray}
				||\mathbf{I} - \mathbf{Q^TQ}|| \leq \frac{\alpha\kappa^2}{1-\beta\kappa^2}.	
			\end{eqnarray}
		}
		\item{Modified Gram-Schmidt is better, but not perfect,
			\begin{eqnarray}
				||\mathbf{I} - \mathbf{Q^TQ}|| \leq \frac{\gamma\kappa}{1-\eta\kappa}	
			\end{eqnarray}
		}
		\item{May need multiple orthogonalization steps}
	\end{itemize}

@@

\begin{frame}{First Brillouin Zone}
	\begin{columns}[c] % align columns
		\begin{column}{.4\textwidth}
			\begin{itemize}
				\item {Excite Ony in the X direction}
				\item {-X virtuals matter due to $\vec{k}' = \vec{k} + \vec{G}$}
			\end{itemize}				
		\end{column}
		\hfill
		\begin{column}{0.6\textwidth}
		    \begin{overprint}
			    \onslide<1>\includegraphics[width=\linewidth]{../images/1stBZ.eps}

\begin{frame}{Orbital Energies}
	\begin{columns}[c] % align columns
		\begin{column}{.4\textwidth}
			\begin{itemize}
				\item {Nk = 57 reproduces the orbital energies reasonably well}
				\item {Worst towards $\Gamma$, better for higher $|\vec{k}|$}
			\end{itemize}				
		\end{column}
		\hfill
		\begin{column}{0.6\textwidth}
		    \begin{overprint}
			    \onslide<1>\includegraphics[width=\linewidth]{../images/energycompare.eps}

\begin{frame}{Matrix Diagonals}
	\begin{columns}[c] % align columns
		\begin{column}{.4\textwidth}
			\begin{itemize}
				\item {Spectrum is Dense}
			\end{itemize}				
		\end{column}
		\hfill
		\begin{column}{0.6\textwidth}
		    \begin{overprint}
			    \onslide<1>\includegraphics[width=\linewidth]{../images/exchist.eps}

@@

{%
\setbeamertemplate{frame footer}{1) http://mathworld.wolfram.com/StrassenFormulas.html}

\begin{frame}{Davidson Scaling}
	\begin{columns}[c] % align columns
		\begin{column}{.4\textwidth}
			\begin{itemize}
				\item {Davidson is Asymptotically quadratic.}
				\item {Full diagonalization is almost cubic.}
				\item {Matrix multiplication is order${}^1$ $log_2(7) \approx 2.807$}
			\end{itemize}				
		\end{column}
		\hfill
		\begin{column}{0.6\textwidth}
		    \begin{overprint}
			    \onslide<1>\includegraphics[width=\linewidth]{../images/dav_vs_exact_scaling.eps}

@@

\begin{frame}{Algorithm Wall time}
	\begin{columns}[c] % align columns
		\begin{column}{.4\textwidth}
			\begin{itemize}
				\item {Roughly quadratic wrt matrix dimension.}
				\item {Outliers correspond to high condition number}
			\end{itemize}				
		\end{column}
		\hfill
		\begin{column}{0.6\textwidth}
		    \includegraphics[width=\linewidth]{../images/overallruntime.eps}
		\end{column}	
	\end{columns}
\end{frame}

\begin{frame}{Efficacy of Davidson's Algorithm}
	\begin{columns}[c] % align columns
		\begin{column}{.4\textwidth}
			\begin{itemize}
				\item {Reproduces Exact result to machine precision in all test cases.}
				\item {Odd spikes are due to approximating circle by squares}
			\end{itemize}				
		\end{column}
		\hfill
		\begin{column}{0.6\textwidth}
		    \begin{overprint}
			    \onslide<1>\includegraphics[width=\linewidth]{../images/dav_vs_exact.eps}

@@

{%
\setbeamertemplate{frame footer}{Bernu, B.; Delyon, F.; Holzmann, M.; Baguet, L. Phys. Rev. B 2011, 84 (11), 115115.}
\begin{frame}{Dependence on r${}_s$}
	Crossover from stable to unstable agrees with previous results.
	\includegraphics[width=.95\linewidth]{../images/MeVsBernu.pdf}
\end{frame}

@@

{%
\setbeamertemplate{frame footer}{}
\begin{frame}{Next Steps}
	\resizebox{\textwidth}{!}{%
	\begin{tabular}{l c c}
		Objective                 & Human Time   & Computer Time    \\
		\midrule
		Other instabilities in 2D & 30 mins & Week \\
		All 3D & $\frac{1}{2}$ Day&  $>$Weeks? Parallel?
	\end{tabular}}
\end{frame}
}

{%
\setbeamertemplate{frame footer}{}
\begin{frame}[standout]
  Questions?
\end{frame}

@@

%{%
%\setbeamertemplate{frame footer}{}
%\begin{frame}[fragile]{Backup slide}
%	Empty Backup
%\end{frame}
%
\end{document}
